### 2017-10-20 code/appendix.Rmd
WIP

---
title: "Appendix"
output: html_notebook
---

```{r}
knitr::opts_chunk(message=FALSE)
```

```{r}
library(nimble)
library(tidyverse)
```

## Stochastic inflation

```{r}

logistic  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }

})

p <- list(r = .4, K = 1, sigma = .2, N = 500)

model <- 
nimbleModel(logistic,
            constants = p,
            inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)

df <- map_dfr(1:10000, 
              function(rep){
                simulate(cmodel)
                data.frame(t = seq_along(cmodel$x), x = cmodel$x)
              },
              .id = "rep")

```

```{r}
(p$K + sqrt(p$K^2 - 8 * p$sigma^2) )/ 2
```

```{r}
df %>% filter(t > 20, x > 0) %>%  summarize(mean(x))
```

```{r}

df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  gather(moment, value, -t) %>%
  ggplot(aes(t,value, col=moment)) + geom_line()

```

------------

# Quasi-cycles

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings]

```{r}

quasicycle  <- nimble::nimbleCode({

  x[1] <- x0
  y[1] <- y0

  for(t in 1:(N-1)){
    mu_x[t] <- x[t] + x[t] * r * (1 - x[t] / K) - b * x[t] * y[t]
    x[t+1] ~ dnorm(mu_x[t], sd = sigma_x)
    mu_y[t] <- y[t] + c * x[t] * y[t] - d * y[t]
    y[t+1] ~ dnorm(mu_y[t], sd = sigma_y)
  }

})

p <- 
  data.frame(
  data.frame(r = .1, K = 5,  b = .1, c = .1, d = .1, N = 500),
  data.frame(sigma_x = c(.00001,0.01),  sigma_y = c(.00001,0.01)))

f <- function(constants){
  model <- compileNimble(nimbleModel(quasicycle, constants = constants, inits = list(x0 = 1, y0 = 1)))
  set.seed(123)
  simulate(model)
  tibble(t = seq_along(model$x), x = model$x, y = model$y, sigma = constants$sigma_x)
}

df <- p %>% rowwise() %>% do(f(.))

```

```{r}
df %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + facet_wrap(~sigma)
```

# Stochastic oscillator

```{r}
p <- list(r = .5, K = 2, Q = 5, H = .38, sigma = .04, a = 0.245, N = 1e4)

f <- function(x) x * p$r * (1 - x / p$K)
g <- function(x) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)

x <- seq(0,2, length.out = 100)
data.frame(x, f = f(x), g = g(x)) %>%
  ggplot(aes(x)) + geom_line(aes(y = f-g)) + 
  geom_hline(aes(yintercept=0), lty=2) +
  coord_cartesian(xlim = c(0,1.5), ylim=c(-.015,.1))

```

```{r}
system.time({
may  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)  - a * x[t] ^ Q / (x[t] ^ Q + H ^ Q)
    y[t+1] ~ dnorm(mu[t], sd = sigma)
    x[t+1] <- max(y[t+1],0)
  }

})

model <- nimbleModel(may,constants = p, inits = list(x0 = 1.2))
cmodel <- compileNimble(model)
set.seed(123)

simulate(cmodel)
df <- tibble(t = seq_along(cmodel$x), x = cmodel$x)
})

df  %>% ggplot(aes(t,x)) + geom_line()
```

# Tipping points

Needs a model with non-zero alternate state!

```{r}

tip  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K) * (x[t] - A) / K - h[t]
    y[t+1] ~ dnorm(mu[t], sd = sigma)
    x[t+1] <- max(y[t+1],0)
  }

})

p <- list(r = .2, K = 1, A = 0.2, sigma = .01, N = 500, h = seq(0,.03,length=500))

model <- nimbleModel(tip,constants = p, inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)

simulate(cmodel)
df <- tibble(t = seq_along(cmodel$x), x = cmodel$x)

```

```{r}
df %>% ggplot(aes(t,x)) + geom_line()
```

# Turing patterns?

------------

```{r}
data(lynx)
lynx <- tibble(year = 1821:1934, lynx = as.numeric(lynx))

lynx %>%
  ggplot(aes(year,lynx)) + geom_point() + geom_line()
```