### 2017-10-21 cover-letter.Rmd

---
title: "Cover Letter"
author: "Carl Boettiger"
date: "10/21/2017"
output: pdf_document
---

- Tim Coulson

Noisy phenomena in population biology.

<!--

We hope that these papers will combine a mix of synthetic reviews of the current state of an area, as well as insights into how the field has developed during the last twenty years, discussion of current important questions, and developments and insights that would significantly advance the field. Although certainly not necessary, if there are any past Ecology Letters papers that have had a profound impact on the development of an area – be it beneficial or not – a brief discussion of that impact would be appropriate. We also hope that articles will be written in a thoughtful style, but also provocative if appropriate.

-->