### 2017-10-23 appendix/appendix.Rmd
WIP

---
title: "Appendix"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r}
library(nimble)
library(hrbrthemes)
library(ggthemes)
library(tidyverse)
ggplot2::theme_set(hrbrthemes::theme_ipsum())
```

------------

# Quasi-cycles

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings1996]

```{r}

quasicycle  <- nimble::nimbleCode({

  x[1] <- x0
  y[1] <- y0

  for(t in 1:(N-1)){
    mu_x[t] <- x[t] + x[t] * r * (1 - x[t] / K) - b * x[t] * y[t]
    x[t+1] ~ dnorm(mu_x[t], sd = sigma_x)
    mu_y[t] <- y[t] + c * x[t] * y[t] - d * y[t]
    y[t+1] ~ dnorm(mu_y[t], sd = sigma_y)
  }

})

p <- 
  data.frame(
  data.frame(r = .1, K = 5,  b = .1, c = .1, d = .1, N = 800),
  data.frame(sigma_x = c(.00001,0.01),  sigma_y = c(.00001,0.01)))

f <- function(constants){
  model <- compileNimble(nimbleModel(quasicycle, constants = constants, inits = list(x0 = 1, y0 = 1)))
  set.seed(123)
  simulate(model)
  tibble(t = seq_along(model$x), x = model$x, y = model$y, sigma = constants$sigma_x)
}

quasicycle_df <- p %>% rowwise() %>% do(f(.))        
write_csv(quasicycle_df, "quasicycles.csv")
```

```{r}
read_csv("quasicycles.csv") %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + facet_wrap(~sigma) + scale_color_ptol()
```

# Stochastic oscillator

May's model

```{r}
p <- list(r = .5, K = 2, Q = 5, H = .38, sigma = .04, a = 0.245, N = 1e4)

f <- function(x) x * p$r * (1 - x / p$K)
g <- function(x) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)

x <- seq(0,2, length.out = 100)
data.frame(x, f = f(x), g = g(x)) %>%
  ggplot(aes(x)) + geom_line(aes(y = f-g)) + 
  geom_hline(aes(yintercept=0), lty=2) +
  coord_cartesian(xlim = c(0,1.5), ylim=c(-.015,.1))

```

```{r}

may  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)  - a * x[t] ^ Q / (x[t] ^ Q + H ^ Q)
    y[t+1] ~ dnorm(mu[t], sd = sigma)
    x[t+1] <- max(y[t+1],0)
  }

})

model <- nimbleModel(may,constants = p, inits = list(x0 = 1.2))
cmodel <- compileNimble(model)
set.seed(123)
simulate(cmodel)

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% write_csv("noisy_switch.csv")
read_csv("noisy_switch.csv")  %>% ggplot(aes(t,x)) + geom_line()
```

## Stochastic inflation

Analytical predicted population size, as a fraction of $K$, is $\tfrac{1}{2} + \tfrac{1}{2} \sqrt{1- 8\sigma^2/K^2}$.  Note this is independent of $r$, and increases with $\sigma/K$, the larger the noise $\sigma_g$ as a fraction of the carrying capacity $K$.  Population sizes can also be inflated relative to their deterministic equilibrium whenever the second derivative is positive (as in the lower equilibrium in the May model of alternative stable states, considered below.)  Stronger nonlinearities (in particular, a larger second derivative at the equilibrium) can drive larger inflationary effects.

```{r}
p <- list(r = .4, K = 1, sigma = .15, N = 500)
equib <- (1 + sqrt(1 - 8 * p$sigma^2/p$K^2) )/ 2

logistic  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }

})

model <- 
nimbleModel(logistic,
            constants = p,
            inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)

df <- map_dfr(1:10000, 
              function(rep){
                simulate(cmodel)
                data.frame(t = seq_along(cmodel$x), x = cmodel$x)
              },
              .id = "rep")
```

```{r}
df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  mutate(equib = equib) %>%
  write_csv("inflation.csv")

```

```{r}
read_csv("inflation.csv") %>% 
  gather(statistic, value, -t) %>%
  ggplot(aes(t,value, col=statistic, lty=statistic)) + 
 # ggplot(aes(t,ave)) +
  geom_line(lwd=1) + 
  scale_color_ptol() + 
  coord_cartesian(xlim=c(100,500), ylim=c(1.2, .5))

```

```{r}
p <- list(r = .5, K = 2.1, Q = 5, H = .4, sigma = .04, a = 0.27, N = 1e4)

f <- function(x) x * p$r * (1 - x / p$K)
g <- function(x) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)

x <- seq(0,2, length.out = 100)
df <- data.frame(x, f = f(x), g = g(x))

ggplot(df, aes(x)) + geom_line(aes(y = f)) + geom_line(aes(y = g), col="red")

ggplot(df, aes(x)) + geom_line(aes(y = f-g)) + 
  geom_hline(aes(yintercept=0), lty=2) +
  coord_cartesian(xlim = c(0,1.5), ylim=c(-.1,.1))  
```

----

# Tipping points

### Configuration for Dai model

Numerically determine / illustrate tipping point location using small noise:

```{r}
max_days <- 1000

z <- y <- x <- numeric(max_days)
z[1] <- y[1] <- x[1] <- 7e4 # 1.7e5
set.seed(123)
for(day in 1:(max_days-1)){
  x[day+1] <- dai(x[day], epsilon = rnorm(1,0, 0.001), DF = 1790)
  y[day+1] <- dai(y[day], epsilon = rnorm(1,0, 0.001), DF=1799)
  z[day+1] <- dai(z[day], epsilon = rnorm(1,0, 0.001), DF = 1800)

}

df <- data.frame(t = 1:max_days, x = x, y = y, z = z) %>%
  gather(series, pop, -t)

ggplot(df, aes(t, pop, col=series)) + 
  geom_line()

```

$$\rho = \frac{1}{n-1} \frac{\sum_{i = 1}^n \left( x_{i,t} - \textrm{E}(x_t) \right)\left( x_{i,t+1} - \textrm{E}(x_{t+1})  \right)}{\sigma_{x_{t}}  \sigma_{x_{t+1}}}$$

```{r}
roll_acor <- function(x, lag = 1, ...){
  x_t1 = lag(x, n = lag)  
  mu_t = roll_mean(x, ...)
  mu_t1 = roll_mean(x_t1, ...)
  s_t = roll_sd(x, ...)
  s_t1 = roll_sd(x_t1, ...)
  acor = roll_mean( (x - mu_t) * (x_t1 - mu_t1) / (s_t * s_t1),...)
  acor
}
```

## Simulation of Dai model:

```{r}
# Stepwise changes
DF <- as.numeric(sapply(seq(0, 2000, length=9), rep, 40))

# continuous linear increase
DF <- seq(1, 2000, length=2000)
tip_time <- 1800
max_days <- length(DF)
y <- numeric(max_days)

set.seed(1111)
y[1] <- 1.76e5
for(day in 1:(max_days-1)){
  y[day+1] <- dai(y[day], DF = DF[day])
}

tip_df <- tibble(t = seq_along(y), x = y)

tip_df %>% 
  mutate(autocorrelation = roll_acor(x, n = 100, fill=NA),
         variance = roll_var(x, n = 100, fill=NA),
         tip_time = tip_time) %>%
  write_csv("tipping.csv")
```

```{r}
tipping <- read_csv("appendix/tipping.csv")

p1 <- tipping %>%
  select(-x) %>%
  gather(series, value, -t, -tip_time) %>%
  ggplot(aes(t, value)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) +
  facet_wrap(~series, ncol = 2, scales="free_y") + theme_ipsum() + 
  labs(title="A: Early warning signs")

p2 <- tipping %>% 
  select(t, population = x, tip_time) %>%
  ggplot(aes(t, population)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) + 
  scale_y_log10() + 
  theme_ipsum()+ labs(title="B: Population density",
                      caption = "Vertical red dashed line indicates tipping point location")

library(cowplot)
ggdraw() +
  draw_plot(p2, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot(p1, x = 0, y = .4, width = 1, height = 0.5)

```