### 2017-10-27 appendixA/appendixA.Rmd
use gridExtra not cowplot

---
title: "Appendix A: Stochastic simulations of noisy phenomena"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r}
library(nimble)
library(RcppRoll)
library(regimeshifts) ## devtools::install_github("cboettig/regimeshifts")
library(adaptivetau)
library(tidyverse)

## plot themes, non-essential
library(ggthemes)
library(gridExtra)
#library(hrbrthemes)
#ggplot2::theme_set(hrbrthemes::theme_ipsum())
ggplot2::theme_set(theme_bw())
colours <- ptol_pal()(2)

```

# Gillespie algorithm

Levins' Patch model as an individual-based birth-death process:

$$\frac{dn}{dt} = \underbrace{c n \left(1 - n/N \right)}_{\textrm{birth rate, } b(n)} - \underbrace{en}_{\textrm{death rate, } d(n)} $$

We simulate this using the Gillespie's exact stochastic simulation algorithm (SSA), as implemented in the `adaptivetau` R package.  We declare the transition events: birth increases the state $n$ to $n+1$, and death $n$ to $n-1$, and describe the rates associated with them:

```{r}
# indicate state variable and the change
transitions = list(c(n = +1), # birth event
                   c(n = -1)) # death event

rateF <- function(state, params, t) {
  c(params$c * state[["n"]] * (1 - state[["n"]] / params$N), # birth rate
    params$e * state[["n"]])                                 # death rate
}

```

We compare the stochastic simulation with identical parameters, $c=1$, $e=0.2$ except for the total number of available sites.  In our "small" system scenario, we have a total of $N=100$ sites, while in the "large" system we have $N=1000$ sites.

The theory predicts an equilibrium population size $\bar n$ at $b(\hat n) = d(\hat n)$, or

$$\left(1 - \tfrac{e}{c}\right) N,$$ 
and a variance of

$$\left. \frac{b(n) + d (n)}{2 d'(n) - b'(n)} \right\rvert_{n = \hat n} = \tfrac{e}{c}N$$

```{r gillespie}
set.seed(1234) # set random number generator seed to be reproducible

e <- 0.2
c <- 1

sim_small <- adaptivetau::ssa.exact(init.values = c(n = 50),
                       transitions, rateF,
                       params = list(c=c, e=e, N=100),
                       tf=30) %>% as_tibble()

sim_large <- adaptivetau::ssa.exact(init.values = c(n = 500),
                       transitions, rateF,
                       params = list(c=c, e=e, N=1000),
                       tf=30) %>% as_tibble()
```

```{r}
theory <- 
list(small = 100,large = 1000) %>%
  map_dfr(function(N){
    data.frame(mean = N * (1-e/c)) %>%
    mutate(plus_sd = mean + sqrt(N * e / c),
           minus_sd = mean - sqrt(N * e / c))
  }, .id = "system_size")
```

```{r}
## Bind together both simulation results and theory results
gillespie <- bind_rows("large"=sim_large, 
                       "small"=sim_small,  
                       .id = "system_size") %>% bind_rows(theory)
write_csv(gillespie, "gillespie.csv")
```

```{r}
read_csv("gillespie.csv", col_types = "cdiddd")  %>% 
  ggplot(aes(x = time)) + 
  geom_hline(aes(yintercept = mean), lty=2) + 
  geom_hline(aes(yintercept = minus_sd), lty=2) + 
  geom_hline(aes(yintercept = plus_sd), lty=2) + 
  geom_line(aes(y = n), col="red") +
  facet_wrap(~system_size, scales = "free_y") + 
  labs(caption = "theoretical predictions for mean and plus/minus one standard \n deviation shown in horizontal black dashed lines.")
```

------------

# Quasi-cycles

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings1996]

```{r}

quasicycle  <- nimble::nimbleCode({

  x[1] <- x0
  y[1] <- y0

  for(t in 1:(N-1)){
    mu_x[t] <- x[t] + x[t] * r * (1 - x[t] / K) - b * x[t] * y[t]
    x[t+1] ~ dnorm(mu_x[t], sd = sigma_x)
    mu_y[t] <- y[t] + c * x[t] * y[t] - d * y[t]
    y[t+1] ~ dnorm(mu_y[t], sd = sigma_y)
  }

})

p <- 
  data.frame(
  data.frame(r = .1, K = 5,  b = .1, c = .1, d = .1, N = 800),
  data.frame(sigma_x = c(.00001,0.01),  sigma_y = c(.00001,0.01)))

f <- function(constants){
  model <- compileNimble(nimbleModel(quasicycle, constants = constants, inits = list(x0 = 1, y0 = 1)))
  set.seed(123)
  simulate(model)
  tibble(t = seq_along(model$x), x = model$x, y = model$y, sigma = constants$sigma_x)
}

quasicycle_df <- p %>% rowwise() %>% do(f(.))        
write_csv(quasicycle_df, "quasicycles.csv")
```

```{r}
read_csv("quasicycles.csv") %>%
  gather(species, pop, -t, -sigma) %>%
  mutate(sigma = as.factor(sigma)) %>%
  mutate(sigma = recode(sigma, "1e-05" = "A.", "0.01" = "B.")) %>%
  ggplot(aes(t, pop, col=species)) +
  geom_line() + 
  facet_wrap(~sigma) + 
  scale_color_ptol()
```

# Stochastic oscillator

May's model:

$$X_{t+1} = X_t + \underbrace{X_t r \left(1 -\frac{X_t}{K} \right)}_{\textrm{Vegitation growth}}  - \underbrace{\frac{a X_t ^ Q}{X_t^ Q + H ^ Q}}_{\textrm{Vegitation consumption}} + \xi_t,$$

```{r}

p <- list(r = .5, K = 2, Q = 5, H = .38, sigma = .04, a = 0.245, N = 1e4)
f <- function(x) x * p$r * (1 - x / p$K)
g <- function(x) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)
x <- seq(0,2, length.out = 100)

data.frame(x, growth = f(x), consumption = g(x)) %>%
  gather(curve, y, -x) %>%
  ggplot(aes(x, y, col=curve)) +
  geom_line(lwd=1) + 
  scale_color_ptol() + 
  labs(x = bquote(X[t]), y = bquote(X[t+1]), subtitle= bquote(bold(B.)))
```

```{r}

may  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)  - a * x[t] ^ Q / (x[t] ^ Q + H ^ Q)
    y[t+1] ~ dnorm(mu[t], sd = sigma)
    x[t+1] <- max(y[t+1],0)
  }

})

model <- nimbleModel(may,constants = p, inits = list(x0 = 1.2))
cmodel <- compileNimble(model)
set.seed(123)
simulate(cmodel)

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% write_csv("noisy_switch.csv")
read_csv("noisy_switch.csv")  %>% ggplot(aes(t,x)) + geom_line()
```

## Stochastic inflation

Analytical predicted population size, as a fraction of $K$, is $\tfrac{1}{2} + \tfrac{1}{2} \sqrt{1- 8\sigma^2/K^2}$.  Note this is independent of $r$, and increases with $\sigma/K$, the larger the noise $\sigma_g$ as a fraction of the carrying capacity $K$.  Population sizes can also be inflated relative to their deterministic equilibrium whenever the second derivative is positive (as in the lower equilibrium in the May model of alternative stable states, considered below.)  Stronger nonlinearities (in particular, a larger second derivative at the equilibrium) can drive larger inflationary effects.

```{r}
p <- list(r = .4, K = 1, sigma = .15, N = 500)
equib <- (1 + sqrt(1 - 8 * p$sigma^2/p$K^2) )/ 2

logistic  <- nimble::nimbleCode({

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }

})

model <- 
nimbleModel(logistic,
            constants = p,
            inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)

df <- map_dfr(1:10000, 
              function(rep){
                simulate(cmodel)
                data.frame(t = seq_along(cmodel$x), x = cmodel$x)
              },
              .id = "rep")
```

```{r}
df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  mutate(equib = equib) %>%
  write_csv("inflation.csv")
```

```{r}
read_csv("inflation.csv") %>% 
  gather(statistic, value, -t) %>%
  ggplot(aes(t,value, col=statistic, lty=statistic)) + 
 # ggplot(aes(t,ave)) +
  geom_line(lwd=1) + 
  scale_color_ptol() + 
  coord_cartesian(xlim=c(100,500), ylim=c(1.2, .5))
```

----

# Tipping points

### Configuration for Dai model

Numerically determine / illustrate tipping point location using small noise:

```{r}
max_days <- 1000

z <- y <- x <- numeric(max_days)
z[1] <- y[1] <- x[1] <- 7e4 # 1.7e5
set.seed(123)
for(day in 1:(max_days-1)){
  x[day+1] <- dai(x[day], epsilon = rnorm(1,0, 0.001), DF = 1790)
  y[day+1] <- dai(y[day], epsilon = rnorm(1,0, 0.001), DF=1799)
  z[day+1] <- dai(z[day], epsilon = rnorm(1,0, 0.001), DF = 1800)

}

df <- data.frame(t = 1:max_days, x = x, y = y, z = z) %>%
  gather(series, pop, -t)

ggplot(df, aes(t, pop, col=series)) + 
  geom_line()

```

$$\rho = \frac{1}{n-1} \frac{\sum_{i = 1}^n \left( x_{i,t} - \textrm{E}(x_t) \right)\left( x_{i,t+1} - \textrm{E}(x_{t+1})  \right)}{\sigma_{x_{t}}  \sigma_{x_{t+1}}}$$

```{r}
roll_acor <- function(x, lag = 1, ...){
  x_t1 = lag(x, n = lag)  
  mu_t = roll_mean(x, ...)
  mu_t1 = roll_mean(x_t1, ...)
  s_t = roll_sd(x, ...)
  s_t1 = roll_sd(x_t1, ...)
  acor = roll_mean( (x - mu_t) * (x_t1 - mu_t1) / (s_t * s_t1),...)
  acor
}
```

## Simulation of Dai model:

```{r}
# Stepwise changes
DF <- as.numeric(sapply(seq(0, 2000, length=9), rep, 40))

# continuous linear increase
DF <- seq(1, 2000, length=2000)
tip_time <- 1800
max_days <- length(DF)
y <- numeric(max_days)

set.seed(1111)
y[1] <- 1.76e5
for(day in 1:(max_days-1)){
  y[day+1] <- dai(y[day], DF = DF[day])
}

tip_df <- tibble(t = seq_along(y), x = y)

tip_df %>% 
  mutate(autocorrelation = roll_acor(x, n = 100, fill=NA),
         variance = roll_var(x, n = 100, fill=NA),
         tip_time = tip_time) %>%
  write_csv("tipping.csv")
```

```{r}
tipping <- read_csv("tipping.csv")

p1 <- tipping %>%
  select(-x) %>%
  gather(series, value, -t, -tip_time) %>%
  ggplot(aes(t, value)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) +
  facet_wrap(~series, ncol = 2, scales="free_y") +
  #theme_ipsum() + 
  labs(title="A: Early warning signs")

p2 <- tipping %>% 
  select(t, population = x, tip_time) %>%
  ggplot(aes(t, population)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) + 
  scale_y_log10() + 
#  theme_ipsum() + 
  labs(title="B: Population density",
              caption = "Vertical red dashed line indicates tipping point location")

gridExtra::grid.arrange(p1,p2, layout_matrix = rbind(c(1,1), c(2,2)))

```